---
title: "a4_task1_dani_sclafani"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
library(here)
library(naniar)
library(janitor)

library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```

## Task 1: Creating an Agglomerative Hierarchical Clustering

For Task 1, use hierarchical clustering by complete linkage to create a dendrogram showing multivariate clustering for water chemistry by site,

To perform hierarchical clustering by site, you’ll want to make a data frame that has a single summary row per site (e.g. based on means from all observations at that site), then calculate the euclidean distance before performing complete linkage agglomerative hierarchical clustering. 


```{r}
#reading in the data and replacing -999.0 with NAs

stream_data <- read.csv(here("data", "sbc_lter_registered_stream_chemistry.csv"))%>%
  replace_with_na_all(condition = ~ .x == -999.0)

stream_tidy <- stream_data %>% 
  group_by(site_code) %>% 
  summarize(mean(nh4_uM, na.rm = TRUE),
            mean(no3_uM, na.rm = TRUE),
            mean(po4_uM, na.rm = TRUE),
            mean(tdn_uM, na.rm = TRUE),
            mean(tdp_uM, na.rm = TRUE), 
            mean(tpc_uM, na.rm = TRUE), 
            mean(tpn_uM, na.rm = TRUE), 
            mean(tpp_uM, na.rm = TRUE), 
            mean(tss_mgperLiter, na.rm = TRUE),
            mean(spec_cond_uSpercm, na.rm = TRUE)) %>% 
  drop_na() %>% 
  rename(nh4 = "mean(nh4_uM, na.rm = TRUE)") %>% 
  rename(no3 = "mean(no3_uM, na.rm = TRUE)") %>% 
  rename(po4 = "mean(po4_uM, na.rm = TRUE)") %>% 
  rename(tdn = "mean(tdn_uM, na.rm = TRUE)") %>% 
  rename(tdp = "mean(tdp_uM, na.rm = TRUE)") %>% 
  rename(tpc = "mean(tpc_uM, na.rm = TRUE)") %>% 
  rename(tpn = "mean(tpn_uM, na.rm = TRUE)") %>% 
  rename(tpp = "mean(tpp_uM, na.rm = TRUE)") %>% 
  rename(tss = "mean(tss_mgperLiter, na.rm = TRUE)") %>% 
  rename(spec = "mean(spec_cond_uSpercm, na.rm = TRUE)") 

```

### Part 1.0 Data Visualizations

```{r}
ggplot(data = stream_tidy, aes(x = nh4, y = no3))+ geom_point()+
  labs(x = "Mean Concentration of NH4 (uM)", y = "Mean Concentration of NO3 (uM)")
```
```{r}
#Second data visualizations
ggplot(data = stream_tidy, aes(x = tdp, y = tss))+ geom_point()+
  labs(x = "Mean Concentration of Total Dissolved Phosphorous (uM)", y = "Total Suspended Solids (per Liter)")
```
```{r}
#3rd data visualizations
ggplot(data = stream_tidy, aes(x = tss, y = spec))+ geom_point()+
  labs(x = "Total Suspended Solids (per Liter)", y = "Specific Conductivity (per cm)")
```

#### need to get this to work but moving on for now
```{r}
#asking R how many clusters it thinks there should be

cluster_est <- NbClust(stream_tidy[2:11], min.nc = 2, max.nc = 10, method = "kmeans")
```





Data Citation: Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174. 
