---
title: "a4_task1_dani_sclafani"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
library(here)
library(naniar)

library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```

## Task 1: Creating an Agglomerative Hierarchical Clustering

For Task 1, use hierarchical clustering by complete linkage to create a dendrogram showing multivariate clustering for water chemistry by site,

To perform hierarchical clustering by site, you’ll want to make a data frame that has a single summary row per site (e.g. based on means from all observations at that site), then calculate the euclidean distance before performing complete linkage agglomerative hierarchical clustering. 


```{r}
#reading in the data

stream_data <- read.csv(here("data", "sbc_lter_registered_stream_chemistry.csv"))

stream_data_tidy <- stream_data %>%
  replace_with_na_all(condition = ~ .x == -999.0)

```




Data Citation: Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174. 
