---
title: "sclafani_dani_a4_task3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message= FALSE)

library(tidyverse)
library(here)
library(janitor)
library(boot)
library(gt)
library(patchwork)
library(broom)
library(nlstools)

```

### Data Wrangling
```{r}
#reading in the dataset I made from the ucsb pdf. created a csv containing 1's for students who experienced exclusionary conduct and 0 for students who did not
conduct_data <- read_csv(here("data", "conduct_data.csv")) %>% clean_names()

#visualizing the data
hist(conduct_data$number)
mean(conduct_data$number)
t.test(conduct_data$number)
```

### Bootstrapping the Sample to find a 95% confidence interval for the proportion of genderqueer individuals who experience exclusionary conduct based on UCSB's 2014 Campus Climate Survey
```{r}
# first need to find a formula that will calculate the proportion of each bootstrapped sample
funct_proportion <- function(x,i) {mean(x[i])}

#create a vector of conduct experiences
conduct_vec <- conduct_data$number

#10k bootstrap samples
conduct_boot_10k <- boot(conduct_vec,
                         statistic = funct_proportion,
                         R = 10000)
```




### Plotting the Bootstrap samples
```{r}
#plot of the original data
plot_og <- ggplot(data= conduct_data, aes(x=number)) + geom_histogram()

#plotting the bootstrap samples, need to make bootstrap a data frame first
conduct_boot_df <- data.frame(bs_proportion = conduct_boot_10k$t)

 ggplot(data= conduct_boot_df, aes(x=bs_proportion)) + geom_histogram(fill = "purple", bins =32, color = "black")+
   labs(x= "Proportion of Nonbinary Genderqueer Students \n who Experience Exclusionary Conduct", y = "Count of Bootstrap Samples",
        title= "Proportions of UCSB Genderqueer/Nonbinary Students \nwho Experienced Exclusionary Conduct") +
   theme_minimal()+
   theme(plot.title = element_text(hjust = 0.5))

```

### Finding the 95% confidence interval for nonbinrary/genderqueer UCSB Students who Experienced Exclusionary Conduct (2014)
```{r}
boot.ci(conduct_boot_10k, conf = 0.95)
```

We expect confidence intervals from 9,500 of our 10,000 bootstrap samples to contain the true proportion of UCSB nonbinrary/genderqueer students who experienced exclusionary conduct according to the 2014 Campus Climate Report. Therefore it is very likely the true proportion of students who experienced exclusionary conduct is within the confidence interval (0.33, 0.77). 






