---
title: "a4_task1_dani_sclafani"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
library(here)
library(naniar)
library(janitor)

library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```

## Task 1: Creating an Agglomerative Hierarchical Clustering

Hierarchical clustering by complete linkage to create a dendrogram showing multivariate clustering for water chemistry by site,

To perform hierarchical clustering by site, first make a data frame that has a single summary row per site (e.g. based on means from all observations at that site), then calculate the euclidean distance before performing complete linkage agglomerative hierarchical clustering. 

#### Data Wrangling
```{r}
#reading in the data and replacing -999.0 with NAs

stream_data <- read.csv(here("data", "sbc_lter_registered_stream_chemistry.csv"))%>%
  replace_with_na_all(condition = ~ .x == -999.0)

stream_tidy <- stream_data %>% 
  group_by(site_code) %>% 
  summarize(mean(nh4_uM, na.rm = TRUE),
            mean(no3_uM, na.rm = TRUE),
            mean(po4_uM, na.rm = TRUE),
            mean(tdn_uM, na.rm = TRUE),
            mean(tdp_uM, na.rm = TRUE), 
            mean(tpc_uM, na.rm = TRUE), 
            mean(tpn_uM, na.rm = TRUE), 
            mean(tpp_uM, na.rm = TRUE), 
            mean(tss_mgperLiter, na.rm = TRUE),
            mean(spec_cond_uSpercm, na.rm = TRUE)) %>% 
  drop_na() %>% 
  rename(nh4 = "mean(nh4_uM, na.rm = TRUE)") %>% 
  rename(no3 = "mean(no3_uM, na.rm = TRUE)") %>% 
  rename(po4 = "mean(po4_uM, na.rm = TRUE)") %>% 
  rename(tdn = "mean(tdn_uM, na.rm = TRUE)") %>% 
  rename(tdp = "mean(tdp_uM, na.rm = TRUE)") %>% 
  rename(tpc = "mean(tpc_uM, na.rm = TRUE)") %>% 
  rename(tpn = "mean(tpn_uM, na.rm = TRUE)") %>% 
  rename(tpp = "mean(tpp_uM, na.rm = TRUE)") %>% 
  rename(tss = "mean(tss_mgperLiter, na.rm = TRUE)") %>% 
  rename(spec = "mean(spec_cond_uSpercm, na.rm = TRUE)") 

```


#### Creating a scaled version of the data
```{r}
#keeping all measurements right now, but maybe i shouldn't? Maybe I only want to keep the measurements of interest
stream_scaled <- stream_tidy %>% 
  select(-site_code) %>% 
  scale()
```


#### Adding back in site_code as rownames 
```{r}
rownames(stream_scaled) <-stream_tidy$site_code
```

#### Finding the Euclidean Distance
```{r}
euc_distance <- dist(stream_scaled, method = "euclidean")

view(euc_distance)
```

#### Performing hierarchical clustering by complete linkage and creating a dendrogram
```{r, fig.cap= "Figure 1.0: Dendrogram of Stream Water Chemistry at 9 Sites"}
#hierarchical clustering with complete linkage
hc_complete <- hclust(euc_distance, method = "complete")

plot(hc_complete, cex = 0.6, hang = -1) 
```


Data Citation: Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174. 
